---
title: "capstone_2"
output: html_document
date: "2023-08-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Capstone II Project Worksheet

The Data:

```{r}
data_nat <- read.csv("C:/Users/jorda/Desktop/Coding_Temple_Work/capstone_2/data/NationalNames.csv")
# View(data_nat)
```

### The Hypothesis:

#### Popularity of a particular name doesn't last longer than about a decade.

Create "Decade" column, categorize values by Decade.

```{r}
data_nat$Decade <- 

data_nat[data_nat$Year >= 1900 & data_nat$Year <= 1909, "Decade"] <- 1900
data_nat[data_nat$Year >= 1910 & data_nat$Year <= 1919, "Decade"] <- 1910
data_nat[data_nat$Year >= 1920 & data_nat$Year <= 1929, "Decade"] <- 1920
data_nat[data_nat$Year >= 1930 & data_nat$Year <= 1939, "Decade"] <- 1930
data_nat[data_nat$Year >= 1940 & data_nat$Year <= 1949, "Decade"] <- 1940
data_nat[data_nat$Year >= 1950 & data_nat$Year <= 1959, "Decade"] <- 1950
data_nat[data_nat$Year >= 1960 & data_nat$Year <= 1969, "Decade"] <- 1960
data_nat[data_nat$Year >= 1970 & data_nat$Year <= 1979, "Decade"] <- 1970
data_nat[data_nat$Year >= 1980 & data_nat$Year <= 1989, "Decade"] <- 1980
data_nat[data_nat$Year >= 1990 & data_nat$Year <= 1999, "Decade"] <- 1990
data_nat[data_nat$Year >= 2000, "Decade"] <- "2000s"
data_nat[data_nat$Year < 1900, "Decade"] <- "pre-1900"

# View(data_nat %>% select("Name","Year","Decade"))
```

Female data first:

```{r}
data_nat_grouped <- group_by(data_nat, Decade, Count)
F_data_nat <- filter(data_nat_grouped, Gender == 'F', Decade != "pre-1900" & Decade != "2000s")
# View(F_data_nat)
```

Now the top 5 Female names of each decade:

```{r}
F_data_sorted <- F_data_nat %>% arrange(desc(Count))
# View(F_data_sorted)

F_nat_1900 <- filter(F_data_sorted, Decade == 1900)
# View(head(F_nat_1900))
F_1900 <- F_nat_1900[!duplicated(F_nat_1900$Name), ]
# View(head(F_1900))

F_nat_1910 <- filter(F_data_sorted, Decade == 1910)
F_1910 <- F_nat_1910[!duplicated(F_nat_1910$Name), ]

F_nat_1920 <- filter(F_data_sorted, Decade == 1920)
F_1920 <- F_nat_1920[!duplicated(F_nat_1920$Name), ]

F_nat_1930 <- filter(F_data_sorted, Decade == 1930)
F_1930 <- F_nat_1930[!duplicated(F_nat_1930$Name), ]

F_nat_1940 <- filter(F_data_sorted, Decade == 1940)
F_1940 <- F_nat_1940[!duplicated(F_nat_1940$Name), ]

F_nat_1950 <- filter(F_data_sorted, Decade == 1950)
F_1950 <- F_nat_1950[!duplicated(F_nat_1950$Name), ]

F_nat_1960 <- filter(F_data_sorted, Decade == 1960)
F_1960 <- F_nat_1960[!duplicated(F_nat_1960$Name), ]

F_nat_1970 <- filter(F_data_sorted, Decade == 1970)
F_1970 <- F_nat_1970[!duplicated(F_nat_1970$Name), ]

F_nat_1980 <- filter(F_data_sorted, Decade == 1980)
F_1980 <- F_nat_1980[!duplicated(F_nat_1980$Name), ]

F_nat_1990 <- filter(F_data_sorted, Decade == 1990)
F_1990 <- F_nat_1990[!duplicated(F_nat_1990$Name), ]

```

Now male name data...

```{r}
M_data_nat <- filter(data_nat_grouped, Gender == 'M', Decade != "pre-1900" & Decade != "2000s")
# View(M_data_nat)
```

Top 5 Male names of each decade:

```{r}
M_data_sorted <- M_data_nat %>% arrange(desc(Count))
# View(M_data_sorted)

M_nat_1900 <- filter(M_data_sorted, Decade == 1900)
# View(head(M_nat_1900))
M_1900 <- M_nat_1900[!duplicated(M_nat_1900$Name), ]
# View(head(M_1900))

M_nat_1910 <- filter(M_data_sorted, Decade == 1910)
M_1910 <- M_nat_1910[!duplicated(M_nat_1910$Name), ]

M_nat_1920 <- filter(M_data_sorted, Decade == 1920)
M_1920 <- M_nat_1920[!duplicated(M_nat_1920$Name), ]

M_nat_1930 <- filter(M_data_sorted, Decade == 1930)
M_1930 <- M_nat_1930[!duplicated(M_nat_1930$Name), ]

M_nat_1940 <- filter(M_data_sorted, Decade == 1940)
M_1940 <- M_nat_1940[!duplicated(M_nat_1940$Name), ]

M_nat_1950 <- filter(M_data_sorted, Decade == 1950)
M_1950 <- M_nat_1950[!duplicated(M_nat_1950$Name), ]

M_nat_1960 <- filter(M_data_sorted, Decade == 1960)
M_1960 <- M_nat_1960[!duplicated(M_nat_1960$Name), ]

M_nat_1970 <- filter(M_data_sorted, Decade == 1970)
M_1970 <- M_nat_1970[!duplicated(M_nat_1970$Name), ]

M_nat_1980 <- filter(M_data_sorted, Decade == 1980)
M_1980 <- M_nat_1980[!duplicated(M_nat_1980$Name), ]

M_nat_1990 <- filter(M_data_sorted, Decade == 1990)
M_1990 <- M_nat_1990[!duplicated(M_nat_1990$Name), ]

```

Female name graphing

```{r}
ggplot(data = head(F_data_sorted,n = 175)) +
  geom_area(mapping = aes(x = Year,y = Count,linetype = Name,fill = Name))

# okay, but not great yet
#ggplot(data = head(F_data_nat,n = 20), mapping = aes(x = Year,y = Count, group = Decade)) +
#  geom_point(mapping = aes(color = Name)) +
#  geom_smooth(se = FALSE)

# nah, not visually clear
# ggplot(data = head(F_nat_1900,n = 20)) +
#   geom_smooth(mapping = aes(x = Year,y = Count, linetype = Name)) +
#   geom_point(mapping = aes(x = Year,y = Count, color = Name))

# just no.
# ggplot(data = head(F_1900,n = 20)) + geom_smooth(mapping = aes(x = Year,y = Count, color = Name))
```

Male name graphing

```{r}
ggplot(data = head(M_data_sorted,n = 175)) +
  geom_area(mapping = aes(x = Year,y = Count,linetype = Name,fill = Name))

# okay, but still not as clear as I'd like
#ggplot(data = head(M_data_nat,n = 20), mapping = aes(x = Year,y = Count, group = Decade)) +
#  geom_point(mapping = aes(color = Name)) +
#  geom_smooth(se = FALSE)

# not clear enough
#ggplot(data = head(M_nat_1900,n = 40)) +
# geom_smooth(mapping = aes(x = Year,y = Count, linetype = Name)) +
# geom_point(mapping = aes(x = Year,y = Count, color = Name))
```

## On to State Data

I'm curious about my home state: Colorado.

```{r}
data_state <- read.csv("C:/Users/jorda/Desktop/Coding_Temple_Work/capstone_2/data/StateNames.csv")
# View(data_state)
```

### Hypothesis:

#### Name trends come and go faster measured in one state than on a national scale. ie: name popularity comes and goes faster in the Colorado data than the national data.

```{r}
# let's get Colorado's data first
CO_data <- filter(data_state, State == 'CO')
# View(CO_data)
```

```{r}
# Now let's make decade groups
CO_data[CO_data$Year >= 1900 & CO_data$Year <= 1909, "Decade"] <- 1900
CO_data[CO_data$Year >= 1910 & CO_data$Year <= 1919, "Decade"] <- 1910
CO_data[CO_data$Year >= 1920 & CO_data$Year <= 1929, "Decade"] <- 1920
CO_data[CO_data$Year >= 1930 & CO_data$Year <= 1939, "Decade"] <- 1930
CO_data[CO_data$Year >= 1940 & CO_data$Year <= 1949, "Decade"] <- 1940
CO_data[CO_data$Year >= 1950 & CO_data$Year <= 1959, "Decade"] <- 1950
CO_data[CO_data$Year >= 1960 & CO_data$Year <= 1969, "Decade"] <- 1960
CO_data[CO_data$Year >= 1970 & CO_data$Year <= 1979, "Decade"] <- 1970
CO_data[CO_data$Year >= 1980 & CO_data$Year <= 1989, "Decade"] <- 1980
CO_data[CO_data$Year >= 1990 & CO_data$Year <= 1999, "Decade"] <- 1990
CO_data[CO_data$Year >= 2000, "Decade"] <- "2000s"
CO_data[CO_data$Year < 1900, "Decade"] <- "pre-1900"

# View(CO_data %>% select("Name","Year","Decade"))
```

```{r}
# Now for female names in CO
CO_data_grouped <- group_by(CO_data, Decade, Count)
F_CO_data <- filter(CO_data_grouped, Gender == 'F', Decade != "pre-1900" & Decade != "2000s")
# View(F_CO_data)

F_CO_sorted <- F_CO_data %>% arrange(desc(Count))
# View(F_CO_sorted)
```

```{r}
# Graphing female names in CO data
ggplot(data = head(F_CO_sorted,n = 150)) +
  geom_area(mapping = aes(x = Year,y = Count,linetype = Name,fill = Name))
```

Now for male names in CO

```{r}
CO_data_grouped <- group_by(CO_data, Decade, Count)
M_CO_data <- filter(CO_data_grouped, Gender == 'M', Decade != "pre-1900" & Decade != "2000s")
# View(M_CO_data)

M_CO_sorted <- M_CO_data %>% arrange(desc(Count))
# View(M_CO_sorted)
```

```{r}
# Graphing male names in CO data
ggplot(data = head(M_CO_sorted,n = 150)) +
  geom_area(mapping = aes(x = Year,y = Count,linetype = Name,fill = Name))
```
